\section*{Declaration} Hereby I declare that this paper is my original authorial work, which
I have worked out on my own. All sources, references, and literature
used or excerpted during elaboration of this work are properly cited
and listed in complete reference to the due source. 

In the course of preparing this thesis, I utilized the following tools based on Artificial Intelligence (AI):
\begin{itemize}
    \item \href{https://www,grammarly.com/}{Grammarly} and \href{https://www.writefull.com/}{Writefull} to enhance my writing style,
    \item \href{https://github.com/features/copilot/}{Github Copilot} to improve my programming efficiency.
\end{itemize}

I declare that I used these tools under the principles of academic integrity. I have reviewed their output and assume full responsibility for it.
\vspace{1cm}
\begin{flushright}
    Tomáš Režňák
\end{flushright}

\vfill\noindent
\textbf{Advisor}: Mgr.~Marek Kadlčík
\cleardoublepage

\section*{Acknowledgments}

I would like to express my sincere gratitude to my advisor, Mgr.~Marek~Kadlčík, for his excellent guidance, invaluable feedback, and, perhaps most importantly, his extraordinary patience throughout the course of my work on this thesis. In addition, I would like to thank those among my friends and family who supported, encouraged, and tolerated me during my studies.

\cleardoublepage

\section*{Abstract}

In this thesis, we present \textsc{svglab}, an open-source Python toolkit for the programmatic manipulation of vector images in the Scalable Vector Graphics (SVG) format. The library provides robust functionality for parsing, modifying, and writing SVG~documents, offering comprehensive support for features defined by the SVG~specification. Next, we conduct a survey of existing software in this domain and demonstrate that our implementation matches or exceeds the capabilities found in even the most feature-rich libraries within the Python ecosystem while addressing some of the common limitations found therein. Finally, we suggest several possible directions for future work in this area.

\vfill

\section*{Keywords}

SVG, SVG manipulation, image processing, vector graphics, affine transformations, parsing, parser, computer graphics, XML

\cleardoublepage
