\chapter{Introduction}\label{chap:introduction}

\gls{svg} has become the de facto standard for vector graphics on the Web and beyond, providing a convenient way to create images that maintain their quality at any size and are often smaller in file size than their raster counterparts.

As with any other format, the need naturally arises to manipulate \gls{svg} documents programmatically, enabling the use of vector graphics in fields such as computer vision and data visualization.

However, despite its widespread use, the Python ecosystem lacks a comprehensive and well-maintained library to manipulate SVG documents. In Section~\ref{sec:other-libs}, we provide a brief overview of existing software and show that only one of the surveyed libraries offers the functionality required to manipulate \gls{svg} documents.

Our contribution is threefold:
\begin{enumerate}
    \item we analyze existing implementations and identify their shortcomings,
    \item we design several novel approaches to solve these shortcomings, and, finally,
    \item we present a new Python library for manipulating \gls{svg} documents, designed to focus on usability, correctness, and customizability.
\end{enumerate}

This paper assumes no prior experience with \gls{svg}. However, we presume that the reader has elementary knowledge of linear algebra, as well as the Python programming language and its ecosystem.

The thesis is structured as follows. Chapter~\ref{chap:preliminaries} covers the preliminaries, including a brief introduction to \gls{svg} and relevant concepts from computer graphics and computer vision. Chapter~\ref{chap:requirements} provides an overview of the requirements for our library. In Chapter~\ref{chap:implementation}, we describe the implementation of the library and its architecture. In Chapter~\ref{chap:evaluation}, we present our efforts to verify the implementation and ensure its correctness, and compare our implementation with existing software in the field. In Chapter~\ref{chap:future-work}, we outline several directions for future work, and finally, we discuss the results in Chapter~\ref{chap:conclusion}.

See Appendix~\ref{appx:source} for a summary of the contents of the digital thesis archive.

We release all our code under the permissive MIT license for the benefit of the community. See Appendix~\ref{appx:installation} for more details and installation instructions.
